\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-07 paper.qmd
Updated Five Gaussian cluster description

The t-SNE, UMAP, and PaCMAP layouts each reveal different aspects of the clustering structure within the dataset. When examining a specific cluster, which appears more like a filled-out pancake in higher dimensions, we observe that t-SNE, UMAP, and PaCMAP all effectively capture global structures. However, there is evidence suggesting that PaCMAP, with its default hyper-parameter settings, struggles to preserve the local structure as effectively as t-SNE and UMAP.

The differences in performance can be attributed to several factors. One key element is the choice of hyper-parameters, particularly the number of neighbors considered during the dimensionality reduction process. This parameter significantly impacts how well local structures are maintained. Additionally, the intrinsic dimensionality of the data, the balance between global and local structures, and issues related to optimization and initialization all contribute to the observed outcomes.

It's important to note that these observations are made with specific hyper-parameter settings. Adjusting these parameters could lead to different results, potentially improving PaCMAP's ability to capture local structures or altering how t-SNE and UMAP handle global and local structures. Thus, while PaCMAP generally balances global and local preservation, careful tuning of its parameters is crucial for achieving the best results in any given dataset.

>>>>

By visualizing the model of different layouts in $4\text{-}D$ reveal different characteristics of each NLDR method. When examining the model of a specific cluster, which appears more like a filled-out shape in tSNE, and UMAP, while the model has a pancake shape in PaCMAP. The reason to have filled-out shaped clusters in tSNE and UMAP is that the ability of capture the local structure. On the other hand, the reason for the pancake shape in PaCMAP is the fail to capture the local structure.

<!-- The tSNE, UMAP, and PaCMAP layouts each reveal different characteristics of the clustering structure within the dataset.  -->

<!-- When examining the model of a specific cluster, which appears more like a filled-out pancake in $4\text{-}D$, we observe that tSNE, UMAP, and PaCMAP all effectively capture global structures. However, there is evidence suggesting that PaCMAP, with its default hyper-parameter settings, struggles to preserve the local structure as effectively as tSNE and UMAP. -->

The differences in performance can be attributed to several factors. One key element is the choice of hyper-parameters, particularly the number of neighbors considered during the dimensionality reduction process. This parameter significantly impacts how well local structures are maintained. Additionally, the balance between global and local structures, and issues related to optimization and initialization all contribute to the observed outcomes.

It's important to note that these layouts are made with default hyper-parameter settings. Adjusting these parameters could lead to different results, potentially improving PaCMAP's ability to capture local structures or altering how tSNE and UMAP handle global and local structures. Thus, while PaCMAP generally balances global and local preservation, careful tuning of its parameters is important for achieving the best results in any given dataset.

### 2024-09-07 paper.tex
Updated Five Gaussian cluster description

The t-SNE, UMAP, and PaCMAP layouts each reveal different aspects of the
clustering structure within the dataset. When examining a specific
cluster, which appears more like a filled-out pancake in higher
dimensions, we observe that t-SNE, UMAP, and PaCMAP all effectively
capture global structures. However, there is evidence suggesting that
PaCMAP, with its default hyper-parameter settings, struggles to preserve
the local structure as effectively as t-SNE and UMAP.

The differences in performance can be attributed to several factors. One
key element is the choice of hyper-parameters, particularly the number
of neighbors considered during the dimensionality reduction process.
This parameter significantly impacts how well local structures are
maintained. Additionally, the intrinsic dimensionality of the data, the
balance between global and local structures, and issues related to
optimization and initialization all contribute to the observed outcomes.

It's important to note that these observations are made with specific
hyper-parameter settings. Adjusting these parameters could lead to
different results, potentially improving PaCMAP's ability to capture
local structures or altering how t-SNE and UMAP handle global and local
structures. Thus, while PaCMAP generally balances global and local
preservation, careful tuning of its parameters is crucial for achieving
the best results in any given dataset.

>>>>

By visualizing the model of different layouts in \(4\text{-}D\) reveal
different characteristics of each NLDR method. When examining the model
of a specific cluster, which appears more like a filled-out shape in
tSNE, and UMAP, while the model has a pancake shape in PaCMAP. The
reason to have filled-out shaped clusters in tSNE and UMAP is that the
ability of capture the local structure. On the other hand, the reason
for the pancake shape in PaCMAP is the fail to capture the local
structure.

The differences in performance can be attributed to several factors. One
key element is the choice of hyper-parameters, particularly the number
of neighbors considered during the dimensionality reduction process.
This parameter significantly impacts how well local structures are
maintained. Additionally, the balance between global and local
structures, and issues related to optimization and initialization all
contribute to the observed outcomes.

It's important to note that these layouts are made with default
hyper-parameter settings. Adjusting these parameters could lead to
different results, potentially improving PaCMAP's ability to capture
local structures or altering how tSNE and UMAP handle global and local
structures. Thus, while PaCMAP generally balances global and local
preservation, careful tuning of its parameters is important for
achieving the best results in any given dataset.
